negation

### 2021-11-29 analysis/exploratory/context_versus_nocontext/rscripts/kl_divergence.Rmd
updated talk for linguae

@@

### Overall
```{r, warning=FALSE, echo=FALSE, message=FALSE}
ggplot(data = d.kls, aes(x=Task, y = kld, fill=Task)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("graphs/denisty_kld_context.pdf",width=6,height=3)
```

### Factors
```{r, warning=FALSE, echo=FALSE, message=FALSE}
agr <- kls.factors %>% 
  group_by(Task, ModalPresent,Wh) %>% 
  summarize(mean_kld = mean(kld), 
            CILow = ci.low(kld), 
            CIHigh = ci.high(kld)) %>%
  mutate(YMin = mean_kld - CILow, 
         YMax = mean_kld + CIHigh) %>%
  drop_na()

>>>>

### Overall
```{r}
agr <- d.kls %>% 
  group_by(Task) %>% 
  summarize(mean_kld = mean(kld), 
            CILow = ci.low(kld), 
            CIHigh = ci.high(kld)) %>%
  mutate(YMin = mean_kld - CILow, 
         YMax = mean_kld + CIHigh) %>%
  drop_na()
dodge = position_dodge(.9)
```

```{r, warning=FALSE, echo=FALSE, message=FALSE}
ggplot(data = agr, aes(x=Task, y = mean_kld, fill=Task)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=dodge, show.legend = FALSE) +
  # ggtitle("Mean rating for 'a' vs. 'every'") +
  xlab("Task") +
  ylab("Mean KL Divergence") +
  guides(fill=FALSE) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,-5,-5),legend.spacing.y = unit(0.001, 'cm')) +
    scale_fill_manual(values=cbPalette) +
    scale_color_manual(values=cbPalette)
# ggsave("../graphs/mean_kld_context.pdf",width=3,height=3)
```

### Factors

@@

```{r, warning=FALSE, echo=FALSE, message=FALSE}
agr <- kls.factors %>% 
  group_by(Task,ModalPresent,Wh) %>% 
  summarize(mean_kld = mean(kld), 
            CILow = ci.low(kld), 
            CIHigh = ci.high(kld)) %>%
  mutate(YMin = mean_kld - CILow, 
         YMax = mean_kld + CIHigh) %>%
  drop_na()

### 2021-11-29 analysis/exploratory/modality/rscripts/graphs.Rmd
updated talk for linguae

# Let's graph overall
```{r agregate for graphing}
agr = d %>%
  # look only at embedded questions
  # filter(VerbLemma != "") %>% 
  group_by(Modal,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()
dodge = position_dodge(.9)
```

```{r graph}
ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position=dodge,stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=dodge, show.legend = FALSE) +
  # ggtitle("Mean rating for 'a' vs. 'every'") +
  facet_wrap(~Modal) +
  xlab("paraphrase") +
  ylab("Mean rating") +
  guides(fill=FALSE) +
  # guides(alpha=guide_legend(title="MatrixNegPresent")) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,-5,-5),legend.spacing.y = unit(0.001, 'cm')) +
    scale_fill_manual(values=cbPalette) +
    scale_color_manual(values=cbPalette)
    # scale_alpha_discrete(range = c(.5,1))
# ggsave("../graphs/overall_negation.pdf",width=6,height=3)

>>>>

```{r look around and refine}
table(d$ModalPresent)
```

The 'd cases seem to all be 'did', not real modals. So we should remove them.
```{r look around and refine}
# look at the 'd cases
t <- d %>% filter(Modal == "'d")
```

```{r remove d}
# remove 'd
d$ModalPresent[d$Modal == "'d"] = "no"
```

```{r rename}
d$Modal[d$Modal == "'ll"] = "will"
d$Modal[d$Modal == "ca"] = "can"
```

```{r look at unlabeled modals}
tt <- d %>% 
  filter((ModalPresent == "yes") & (Modal == ""))

table(tt$Finite)
```

```{r recode to nonfinite}
d$Modal[d$Modal == ""] = "nonfinite"
```

# Let's graph overall
```{r agregate for graphing}
agr = d %>%
  filter(ModalPresent == "yes") %>%
  group_by(Modal,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()
dodge = position_dodge(.9)
```

```{r graph}
ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position=dodge,stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=dodge, show.legend = FALSE) +
  # ggtitle("Mean rating for 'a' vs. 'every'") +
  facet_wrap(~Modal) +
  xlab("paraphrase") +
  ylab("Mean rating") +
  guides(fill=FALSE) +
  # guides(alpha=guide_legend(title="MatrixNegPresent")) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,-5,-5),legend.spacing.y = unit(0.001, 'cm')) +
    scale_fill_manual(values=cbPalette) +
    scale_color_manual(values=cbPalette)
    # scale_alpha_discrete(range = c(.5,1))
ggsave("../graphs/overall_modals.pdf",width=6,height=4)

### 2021-11-29 analysis/exploratory/negation/rscripts/negation.Rmd
updated talk for linguae

```{r bring in the negation coding}
corp = read.table("../../../../corpus/results/swbd.tab",sep="\t",header=T,quote="")
names(corp)
corp <- corp[,c("Item_ID","MatrixNegPresent","MatrixNegation","EmbeddedNegPresent","EmbeddedNegation")]
names(corp)[names(corp) == "Item_ID"] <- "tgrep_id"
table(d$MatrixNegPresent)
```

@@

```{r recode the empty cells as 'no'}
d$MatrixNegPresent[d$MatrixNegPresent != "yes"] = "no"
```

@@

@@

```{r graph}
ggplot(agr,aes(x=paraphrase, y=mean_rating, alpha=MatrixNegPresent, fill=paraphrase)) +
  geom_bar(position=dodge,stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=dodge, show.legend = FALSE) +
  # ggtitle("Mean rating for 'a' vs. 'every'") +
  xlab("paraphrase") +
  ylab("Mean rating") +
  guides(fill=FALSE) +
  guides(alpha=guide_legend(title="MatrixNegPresent")) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,-5,-5),legend.spacing.y = unit(0.001, 'cm')) +
    scale_fill_manual(values=cbPalette) +
    scale_color_manual(values=cbPalette) +
    scale_alpha_discrete(range = c(.5,1))
# ggsave("../graphs/overall_negation.pdf",width=6,height=3)

@@

ggsave("../graphs/overall_negation_eq.pdf",width=3,height=6)

@@

@@

ggsave("../graphs/modxwh_negation_eq.pdf",width=3,height=6)
```

@@

## Look at specific verbs
```{r aggregate and subset to specific verbs}
agr = d %>%
  # Most frequent
  filter(VerbLemma %in% c("know","see","wonder","understand","be","understand","surprise","predict","tell","figure","say","think","sure","learn","remember","forget")) %>%
  group_by(paraphrase,VerbLemma,MatrixNegPresent) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

@@

>>>>

```{r bring in the negation coding}
corp = read.table("../../../../corpus/results/swbd.tab",sep="\t",header=T,quote="")
names(corp)
corp <- corp[,c("Item_ID","MatrixNegPresent","MatrixNegation","EmbeddedNegPresent","EmbeddedNegation")]
names(corp)[names(corp) == "Item_ID"] <- "tgrep_id"
```

@@

```{r recode the empty cells as 'no'}
table(d$MatrixNegPresent)
d$MatrixNegPresent[d$MatrixNegPresent != "yes"] = "no"
```

@@

```{r}
# Make everything into a factor
d$workerid = as.factor(d$workerid)
d$tgrep_id = as.factor(d$tgrep_id)
d$paraphrase = as.factor(d$paraphrase)
d$MatrixNegPresent = as.factor(d$MatrixNegPresent)
```

```{r}
d_eq_msma <- d %>% 
  filter(paraphrase %in% c("a","every"),
         VerbLemma != "" ) %>% 
  mutate(paraphrase.dc = recode(paraphrase, "a" = 1, "every" = 0) ) %>%
  mutate(cParaphrase = as.numeric(paraphrase.dc)-mean(as.numeric(paraphrase.dc))) %>%
  # mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  mutate(cNegation = as.numeric(MatrixNegPresent)-mean(as.numeric(MatrixNegPresent))) %>%
  # mutate(cTask = as.numeric(Task) - mean(as.numeric(Task))) %>%
  droplevels()
nrow(d_eq_msma)
```

# Does negation significantly affect paraphrase?
- i.e., are MS readings really only licesnsed with negatuion?
- answer: not significant.

```{r model}
# m <- lmer(normed_rating ~ cParaphrase*cNegation + (1+cParaphrase|tgrep_id) + (1+cParaphrase+cNegation|workerid), data = d_eq_msma)
# saveRDS(m, "../models/negation.rds")

m <- readRDS("../models/negation.rds")
summary(m)
```

@@

```{r graph}
ggplot(agr,aes(x=paraphrase, y=mean_rating, alpha=MatrixNegPresent, fill=paraphrase)) +
  geom_bar(position=dodge,stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=dodge, show.legend = FALSE) +
  # ggtitle("Mean rating for 'a' vs. 'every'") +
  xlab("paraphrase") +
  ylab("Mean rating") +
  guides(fill=FALSE) +
  guides(alpha=guide_legend(title="MatrixNegPresent")) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,-5,-5),legend.spacing.y = unit(0.001, 'cm')) +
    scale_fill_manual(values=cbPalette) +
    scale_color_manual(values=cbPalette) +
    scale_alpha_discrete(range = c(.5,1))
# ggsave("../graphs/overall_negation.pdf",width=3,height=3)

@@

# ggsave("../graphs/overall_negation_eq.pdf",width=3,height=3)

@@

## Task and negation

```{r aggregate for taskxnegation}
agr = d %>%
  filter(VerbLemma != "") %>% 
  group_by(paraphrase,Task,MatrixNegPresent) %>%
  summarize(mean_rating = mean(normed_rating), 
            CILow = ci.low(normed_rating), 
            CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, 
         YMax = mean_rating + CIHigh) %>%
  drop_na()
```

```{r graph task x negation}
ggplot(agr,aes(x=paraphrase, y=mean_rating, alpha=MatrixNegPresent, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=position_dodge(0.9)) +
  facet_wrap(~Task) +
  xlab("Paraphrase") +
  ylab("Mean rating") +
  guides(fill=FALSE) +
  guides(alpha=guide_legend(title="MatrixNegPresent")) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,-5,-5),legend.spacing.y = unit(0.001, 'cm')) +
    scale_fill_manual(values=cbPalette) +
    scale_color_manual(values=cbPalette) +
    scale_alpha_discrete(range = c(.5,1))

# ggsave("../graphs/task_negation_eq.pdf",width=4,height=3)
```

@@

# ggsave("../graphs/modxwh_negation_eq.pdf",width=3,height=6)
```

@@

## Look at specific verbs
```{r aggregate and subset to specific verbs}
agr = d %>%
  # Most frequent
  filter(VerbLemma %in% c("know","see","wonder","be","understand","surprise","predict","tell","figure","say","think","sure","learn","remember","forget")) %>%
  group_by(paraphrase,VerbLemma,MatrixNegPresent) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()
```

@@

# ggsave("../graphs/matrixverbs_negation_theoretical.pdf",width=10,height=10)
```
